---
title: "TCGA clinical mask"
author: "Andrew"
date: "5/30/2020"
output: html_document
---

```{r setup, include=FALSE}
library(tidyverse)
library(magrittr)
library(here)
library(data.table)
library(ROMOPOmics)
library(knitr)
library(scales)
library(lubridate)
knitr::opts_chunk$set(echo = FALSE)
mask_dir  <- file.path(here(),"masks")
data_file <- file.path(here(),"data/tcga/brca_clinical.csv")
```

#The Cancer Genome Atlas ([TCGA](https://www.cancer.gov/about-nci/organization/ccg/research/structural-genomics/tcga))
##Clinical data mask.

- Add gender value.

- Convert age, death, and date of submission to birthday, death date, and date of submission in datetime format.

- Add means of specifying multiple observations, treatments, etc. at run time. (input has multiple observation values, have ROMOPOmics add these to new columns)

```{r loadDataTable}
tib   <- fread(data_file,header = TRUE) %>%
          as_tibble() %>%
          mutate_all(function(x) ifelse(x=="",NA,x))
```

> File name: `r data_file`

This data table includes **`r prettyNum(ncol(tib) - 1,big.mark=",")` samples** and **`r prettyNum(nrow(tib) - 1,big.mark=",")` metrics**. For the sake of this demo, we assume that the user only wants to use a subset of data included in the TCGA data set. For that reason, I opted to do the following:

###1. Split up samples by number (start with "patient_samples_sample", followed by number OR nothing for 1) into multiple replicates.
```{r parseTable}
tib %>% 
  mutate(field="",table="") %>%
  rename(alias=column_name,example1=V1,example2=V2) %>%
  select(-starts_with("V")) %>%
  write.table(file = file.path(mask_dir,"brca_clinical_mask_blank.tsv"),sep = "\t",row.names = FALSE,col.names=TRUE,quote = FALSE)
```