---
title: 'Example: T-cell activation'
author: "Andrew Clugston"
date: "January 9, 2020"
output: html_document
---

```{r setup, include=FALSE}
library(GenomicRanges)
library(dplyr)
library(knitr)
library(kableExtra)
library(tidyverse)
library(TxDb.Hsapiens.UCSC.hg19.knownGene)
library(scales)
library(data.table)
knitr::opts_chunk$set(echo = FALSE)
source("/pylon5/brz3a1p/codeathon/Chroma_T-Cell/Codes/dir_structure.R")
base_dir<- file.path(dirs$code,"example_analysis")
filter  <- dplyr::filter
select  <- dplyr::select
mutate  <- dplyr::mutate
arrange <- dplyr::arrange
rename  <- dplyr::rename
melt    <- reshape2::melt
```

#Example analysis: **Are there any genes who's promoters become accessible or inaccessible following T-cell activation?**

Question would be best answered with differential enrichment analyses, but we can use BED files of accessible regions to provide a simple binary answer (promoter is open/closed) for each time point measured. We specify a query for:

`Human` `CD4+ T-cells` in `normal donors` at `0,2,4,` and `6 hours post activation`

`This section should be replaced with the R-version of the query function, whatever we end up with.`
```{r loadDataSQL}
#m_test <- read.csv("/pylon5/brz3a1p/codeathon/Chroma_T-Cell/OMOP_tables/molecular_test.txt")
```


```{r loadDataManual}
file_table  <- file.path(dirs$code,"GSE60682_details.tsv") %>%
                read.table(sep="\t",header = TRUE,stringsAsFactors = FALSE) %>%
                as_tibble() %>%
                filter(treatment=="T cell activation")
tst_table   <- file_table %>% select(file_name,sample_name,cell.type,treatment,time.point,local_file_name)
#kable(tst_file)

#Get promoters.
proms       <- promoters(TxDb.Hsapiens.UCSC.hg19.knownGene)

#Read BED files.
read_bed_file   <- function(bed_file_name=tst_table$local_file_name[1]){
  base_col_names<- c("seqnames","start","end","name","score","strand")
  tib <- as_tibble(fread(bed_file_name,sep="\t"))
  if(ncol(tib) > length(base_col_names)){
    base_col_names  <- c(base_col_names,paste0("col.",seq(from=1,to=ncol(tib)-length(base_col_names))))
  }
  colnames(tib)   <- base_col_names
  return(makeGRangesFromDataFrame(tib,keep.extra.columns = TRUE))
}
beds              <- lapply(tst_table$local_file_name,read_bed_file)
names(beds)  <- tst_table$file_name
```

##Promoters that change accessibility on T-cell activation.
```{r prepareTest,fig.height=5,fig.width=7}
test_overlaps   <- function(gr_quer=beds[[1]],gr_subj=proms){
  olaps <- findOverlaps(query = gr_quer,subject = gr_subj)
  out_vec <- rep(FALSE,length(gr_subj))
  out_vec[subjectHits(olaps)] <- TRUE
  return(out_vec)
}
olap_tib <- lapply(beds,test_overlaps) %>%
              do.call(cbind,.) %>%
              as_tibble() %>%
              mutate(tx_name = proms$tx_name) %>%
              select(tx_name,everything())
repd_tib_long <- olap_tib %>%
                  melt(id.vars="tx_name",
                       value.name = "accessible",
                       variable.name = "sample") %>%
                  mutate(time = str_match(sample,pattern="([:digit:])hrs.+$")[,2],
                         rep = str_match(sample,pattern="Rep([:digit:]$)")[,2]) %>%
                  as_tibble() %>%
                  group_by(tx_name,time) %>%
                  summarize(accessible=all(accessible)) %>%
                  ungroup()
repd_tib  <- repd_tib_long %>% 
              spread(key=time,value=accessible) %>%
              mutate(change = case_when(`0`==FALSE & (`1`==TRUE & `2`==TRUE & `4`==TRUE) ~ "Opening",
                                        `0`==TRUE & (`1`==FALSE & `2`==FALSE & `4`==FALSE) ~ "Closing",
                                        TRUE ~ "Uncertain")) %>%
              filter(change != "Uncertain")
```
```{r plotPeakNumbers}
tib <- repd_tib_long  %>%
         group_by(tx_name,time) %>%
         summarize(accessible=all(accessible)) %>%
         ungroup() %>% group_by(accessible,time) %>%
         tally() %>%
         filter(accessible)
  
ggplot(tib,aes(x=as.factor(time),y=n,fill=as.factor(time))) +
       geom_bar(stat="identity",color="black") +
       scale_y_continuous(name="Accessible promotersr",expand=expand_scale(mult=c(0,0.1)),label=comma) +
       scale_x_discrete(name="Hours post-activation") +
       ggtitle("Accessible T-cell promoters following T-cell activation",
               subtitle="Replicated promoters only.") +
       theme(panel.background = element_blank(),
             panel.border = element_rect(size=0.5,color="black",fill=NA),
             panel.grid = element_blank(),
             panel.grid.major.y = element_line(size=0.5,color="lightgray"),
             legend.position="none",
             plot.subtitle = element_text(hjust=0,face="italic"),
             axis.text = element_text(size=15),
             axis.title=element_text(size=20))
ggsave(filename = file.path(base_dir,"peak_count_plot.png"),device = "png",width = 6,height = 4,units = "in",dpi=300)
```
```{r tallyChanges}
repd_tib %>%
  group_by(change) %>%
  tally() %>%
  kable()
```

##Transcripts with changing promoters.
```{r exampleGenes}
tib  <- repd_tib %>%
          select(tx_name,change) %>%
          dplyr::rename(`Transcript ID`=tx_name,
                 Change=change) %>%
          group_by(Change) %>%
          group_split()

cbind(rbind(head(tib[[1]]),c("   ...","  ...")),
      rbind(head(tib[[2]]),c("   ...","  ..."))) %>%
  kable() %>% kable_styling(full_width=FALSE)
```




##PCA of selected samples.
```{r atacPCA}
beds
```